---
title: "Introduction to R"
author: "David Gerard"
date: "`r Sys.Date()`"
output:  
  html_document:
    toc: true
    toc_depth: 4
urlcolor: "blue"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

- R is a statistical programming language designed to analyze data.

- This is not an R course. But you need to know some tools to summarize/plot/model data.

- R is free, widely used, more generally applicable (beyond linear regression), and eases reproducibility. So this is what we will use.

- Python would have been a good choice too, but you can learn that in the machine learning course (STAT 427/627).

# Loading in Data

- Use `read_csv()` from the `{readr}` package to load tabular data into R.

    ```{r}
    library(readr)
    ```

- For other file formats, you need to use other functions, such as `read_tsv()`, `read_table()`, `read_fwf()`, etc. I will try to make sure `read_csv()` works for all datasets in this course.

# Data Frames

```{r, message = FALSE}
library(dplyr)
```

- Use the `mutate()` function from the `{dplyr}` package to make variable transformations.
    
- Use `glimpse()` to get a brief look at the data frame.

- Use `View()` to see a spreadsheet of the data frame (never put this in an R Markdown file). Note the capital "V".

- Use `rename()` to rename variables.

- Use `filter()` to remove rows.

# Plotting

- In this class, we will use the `qplot()` function from the `{ggplot2}` package for all plotting.

    ```{r}
    library(ggplot2)
    ```

- `{ggplot2}` has much more functionality than the `qplot()` function that you can learn about in STAT 412/612.

# Things I expect you to know off the top of your head:

- `help()`: Open help file.
- `install.packages()`: Install an external R package.
- `library()`: Load the functions of an external R package so you can use them.
- `<-`: Variable assignment.
- `+`, `-`, `/`, `*`: Arithmetic operations.
- `^`: Powers.
- `sqrt()`: Square root.
- `log()`: Log (base e).
- `$`: Extracting a variable from a data frame.
- `View()`: Look at a spreadsheet of data.
- `head()`: See first six elements.
- From `{readr}`:
  - `read_csv()`: Loading in tabular data.
- From `{dplyr}`:
  - `glimpse()`: Look at a data frame.
  - `mutate()`: Variable transformation.
  - `rename()`: Variable renaming.
- From `{ggplot2}`
  - `qplot()`: Make a histogram or scatterplot.
